---
title: "Logistic Regression - Predicting Diabetes Diagnosis"
author: "Lealena Befikadu"
subtitle: "4/26/2020"
output: html_document
---
## Data Cleaning/Preparation
Before I analyzed the "diabetes.csv" dataset, I had to do some clean. I first called the libraries needed for my analyzation. I, then, proceeded to set my directory and import the dataset into the variable of my choice. After, I created a variable called height.si and converted the height variable to SI units by multiplying the height variable by 0.0254 into that new variable. I also created a variable called weight.si and converted the weight variable to SI units by multiplying the weight variable by  0.453592 into that new variable. I, then, took those two variables and created a "BMI" variable and put the formula: weight.si/height.si^2, into that variable. Next, I created a categorical BMI variable named bmi_categorised and set conditions to the BMI variable. Those conditions were: bmi < 18.5 was set to equal "underweight", bmi >= 18.5 & bmi <= 25 was set to equal "normal", bmi > 25 & bmi <= 30 was set to equal "overweight", bmi > 30 was set to equal  "obese", and all other values were set to NA. Lastly, I recoded the insurance variable and the smoking variable to reflect their value labels.
```{r warning=FALSE, message=FALSE, results="hide"}
# insert import & data cleaning code here
#set libraries
library(dplyr)
library(ggplot2)
library(tidyverse)
#set directory
setwd("/Users/lealenabefikadu/Downloads")   
#opened file and set it equal to a variable
diabete <- read.csv("diabetes.csv", header = TRUE)
#created new varibale for height with coversion
height.si <- diabete$height*0.0254
#created new varibale for weight with coversion
weight.si <- diabete$weight*0.453592
#created BMI with calculation
BMI <- weight.si/height.si^2
#recode BMI
bmi_categorised<- ifelse(BMI < 18.5,"underweight",
                         ifelse(BMI >= 18.5 & BMI <= 25,"normal",
                                ifelse(BMI > 25 & BMI <= 30,"overweight",
                                       ifelse(BMI > 30,"obese",
                                              NA))))

#recode insurance
diabete$insurance<- ifelse(diabete$insurance==0, "none", 
                            ifelse(diabete$insurance==1,"government", 
                                   ifelse(diabete$insurance==2, "private", NA)))
  

#recode smoking
diabete$smoking<- ifelse(diabete$smoking==1, "current", 
                          ifelse(diabete$smoking==2, "never",
                                 ifelse(diabete$smoking==3, "ex-smoker", NA))) 
```


## Exploratory Data Analysis
While analyzing the "diabetes.csv" dataset, I found there to be 403 observations. The percentage of individuals who are obese in the dataset is 38.3%. The percentage of individuals who are female in the dataset is 58.1%. The percentage of individuals who have diabetes in the dataset is 15.4%. The average age for individuals in the dataset is 46.9 (47). The average BMI for individuals in the dataset is 28.8. 
```{r warning=FALSE, message=FALSE, results="hide"}
# insert exploratory data analysis code here
#num of observations
nrow(diabete)
#percentage of observations for bmi
round(prop.table(table(bmi_categorised))*100, digit = 1)
#percentage of observations for gender
round(prop.table(table(diabete$gender))*100, digit = 1)
#percentage of observations for diabetes
round(prop.table(table(diabete$dm))*100, digit = 1)
#average age
mean(diabete$age)
#average BMI
mean(BMI, na.rm=T)
```


## Data Visualization
For Figure 1, I created a bar chart of insurance. It showed that most individuals in the dataset have government insurance. 
For Figure 2, I created a bar chart of smoking. It showed that most individuals in the dataset never smoked.
For Figure 3, I created a scatter plot of the relationship between BMI and cholesterol. It showed a slightly positive relationship between BMI and cholesterol level, meaning as BMI increases, cholesterol levels increase as well. 
For Figure 4, I created a histogram of the individuals' ages. It showed that most individuals in the dataset are between the ages of 30-50 years old.


**Figure 1**
```{r warning=FALSE, message=FALSE}
# insert data visualization code here
#bar chart for labeled insurance
ggplot(data=diabete) +
  geom_bar(mapping = aes(x=insurance, fill = insurance)) +
  labs(title="Figure 1: Plot of Insurance Status",x="Insurance Status", fill= "Insurance Status",y="Counts")
```
**Figure 2**
```{r warning=FALSE, message=FALSE}
# insert data visualization code here
#bar chart for smoking
ggplot(data=diabete) +
  geom_bar(mapping = aes(x=smoking, fill = smoking)) +
  labs(title="Figure 2: Plot of Smoking Status",x="Smoking Status", fill="Smoking Status",y="Counts")
```
**Figure 3**
```{r warning=FALSE, message=FALSE}
# insert data visualization code here
#scatter plot for BMI and cholesterol
ggplot(diabete, aes(x = BMI, y = chol)) +
            geom_point() +
            geom_smooth(method="lm")+
  labs(title="Figure 3: Relationship between BMI and Cholesterol",x="Body Mass Index", y = "Cholesterol Levels")
```
**Figure 4**
```{r warning=FALSE, message=FALSE}
# insert data visualization code here
#histogram for age
ggplot(data=diabete) + geom_histogram(mapping = aes(x=age), fill="blue", color="green", binwidth = 5) + labs(title="Figure 4: Plot of Ages",x="Age of Individuals",y="Counts")
```


## Logistic Regression Results
The results of the logistic regression showed that for a 1-year increase in age, the logs of having diabetes increases by 0.055454, holding the other independent variables constant. For a l-unit increase in body mass index, the logs of having diabetes increases by 0.073879, holding the other independent variables constant. The log odds of having diabetes between a male and a female is 0.244852, holding the other independent variables constant.
```{r warning=FALSE, message=FALSE, results="hide"}
# insert logistic regression code
logD <- glm(dm ~ age + BMI + gender, data = diabete, family = "binomial")
summary(logD)

```

